---
title: "My Mental Model"
author: "Aaron"
date: "5/29/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.align = "center")
if(!require("pacman"))install.packages("pacman")
pacman::p_load("DiagrammeR")

```

## Why

> *An article about computational results is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.*
> Clearbout & Karrenbach (1992)

CLAERBOUT, J. AND KARRENBACH, M. (1992). Electronic documents give reproducible research a new meaning.
In: Proceedings of the 62nd Annual International Meeting of the Society of Exploration Geophysics, pp. 601–604


## About Philosophie

* the label "Reproducible" just say that you adhere to a lose set of standards in this communication
* there is no real authority, which set these standards
* instead it is the mix of OpenSource and OpenResearch Communities, which adhere to a certain philosophies
* adhering to this philosphie cost a bit of time, there is no free lunch
* but the payoff for you is considerable
* even though the tools change rapidly from day to day this philosophies don't

## Goals for Today

* therefore I'll try to give you the gist of some principles following this philosophies
* **Disclaimer** this "gist" is prepared for you in my highly opinionated way, if you want a reason why I do something in that way, I'll give mostly answers I invent for the purpose of answering the question
* apart from Git(Hub) I don't intend to teach you the tools (today)
* instead I'll give you starting points for tools which aim to make the realization of this principles easy
* learning of the concepts should precedent the tooling
* Rule of Thumb: **Any Project where you work longer on then two hours, has a positive Return of Invest. So try to adhere to these standards if you think you'll need more then an hour, becouse you'll need at least two.**

## Communikation is Key

* Research in general is a lot about communication, therefore ideal communication is a good mental model for reproducible research
* you want to communicate something so precisely that everyone (who knows what he is doing) can reproduce it

# Making Bread

## Expectations

* imagine you want to tell someone the recipe for making bread
* you don't start a recipe with:
    * Step one: You go to the Kitchen
    * Step two: You measure the kitchen to buy a oven."
* instead you (and the person you're communicating with) expect certain things
* the key is to exploit these assumptions/expectations to archive efficient communication
* **BUT**, better be to explicit, then to assume to much

## Structure

* All Coockbooks adhere to the same standart for a reason
* they have an introduction, for what audience, what type of recipies -> `readme.md`
* They start with tools, what do I need to make that -> `library` at the start of an script
* They describe basic technics -> `functions`
* They have common boxes where you look first if you want to cook something -> `folder structure`
* they don't repeat themself, for example a baking book explains how to make short pastry only once and link it -> `reuse of scripts`

# Communication with who?

## Other

```{r}
grViz("
digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, fontsize = 10]

  # several 'node' statements
  node [shape = circle,
        fixedsize = true,
        width = 0.9] // sets as circles
  You; Other

  # several 'edge' statements
  You -> Other

  subgraph {
  rank = same; You; Other;
  }
}
")
```

## Future You

```{r}
grViz("
digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, fontsize = 10]

  # several 'node' statements
  node [shape = circle,
        fixedsize = true,
        width = 0.9] // sets as circles
  You; Other; 'Future \n You'

  # several 'edge' statements
  You -> Other
  You -> 'Future \n You'
  subgraph {
  rank = same; You; Other;
  }
}
")
```

## In all directions

```{r}
grViz("
digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, fontsize = 10]

  # several 'node' statements
  node [shape = circle,
        fixedsize = true,
        width = 0.9] // sets as circles
  You; Other; 'Future \n You'

  # several 'edge' statements
  You -> Other
  You -> 'Future \n You'
  'Other' -> 'Future \n You'
  'Other' -> 'Future \n Other'
  You -> 'Future \n Other'
  'Future \n You' -> 'Future \n Other'
  'Future \n Other' -> 'Future \n You'

  subgraph {
  rank = same; You; Other;
  }
  subgraph {
  rank = same; 'Future \n You'; 'Future \n Other';
  }
}
")
```

# Principles

## Mapping

* every empirical paper is essentially a mapping from data to an digestible output

```{r}
grViz("
digraph boxes_and_circles {
  rankdir=LR

  # a 'graph' statement
  graph [overlap = true, fontsize = 10]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]
  script

  node [shape = circle,
        fixedsize = true,
        width = 0.9] // sets as circles
  data; output;

  # several 'edge' statements
  data -> script -> output
}
")
```

## Two Principles

```{r}
grViz("
digraph boxes_and_circles {
  rankdir=TD

  # a 'graph' statement
  graph [overlap = true, fontsize = 10]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]
  script

  node [shape = circle,
        fixedsize = true,
        width = .8] // sets as circles
  data; output;

  node [shape = diamond,
        fixedsize = true,
        width = 1.6] // sets as circles
  portable; transparent; 

  # several 'edge' statements
  data -> script -> output
  data -> transparent
  output -> portable

  subgraph {
  rank = same; data; script; output;
  }
}
")
```

## Reproducibility

```{r}
grViz("
digraph boxes_and_circles {
  rankdir=TD

  # a 'graph' statement
  graph [overlap = true, fontsize = 10]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]
  script

  node [shape = circle,
        fixedsize = true,
        width = .8] // sets as circles
  data; output;

  node [shape = diamond,
        fixedsize = true,
        width = 1.6] // sets as circles
  portable; transparent; 

  # several 'edge' statements
  data -> script -> output
  script -> {portable; transparent}
  data -> transparent
  output -> portable

  subgraph {
  rank = same; data; script; output;
  }
}
")
```

## Portability

* things should work regardless of the device and operating system
* they should be well ordered
    * everything should be in one place
    * available via network
* publicly available
    * free
    * not bound to restrictive license agreements
* adherence to standards
    * plain text
    * ISO norm
    * style guides

## Transparency

**everthing should be documentated**

* this especially "easy" via programming languages
* everything should be scripted
* what is done to the data should be clearly, concisely and digestible communicated
* not only the end result, but also the steps in between

## Human & Machine

```{r}
grViz("
digraph boxes_and_circles {
  rankdir=TD

  # a 'graph' statement
  graph [overlap = true, fontsize = 10]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]
  script

  node [shape = circle,
        fixedsize = true,
        width = .8] // sets as circles
  data; output; human; machine;

  node [shape = diamond,
        fixedsize = true,
        width = 1.6] // sets as circles
  portable; transparent; 

  # several 'edge' statements
  data -> script -> output
  script -> {portable; transparent}
  portable -> {machine; human}
  transparent -> {machine; human}
  data -> transparent
  output -> portable

  subgraph {
  rank = same; data; script; output;
  }
}
")
```

# Wisdome from Software Engineering

## DRY

**Don't repeat yourself**

> The rule: If you repat something more then three times, write a function, use apply or loop or write a script to automatise what you're doing.

## KISS

**Keep it simple stupid**

```
a <- 1:5
b <- 2:6
```

```
out <- vector("numeric", length(a))
for(i in seq_along(a)){
  out[i] <- a[i] + b[i]
}
```

or

```
a+b
```

## DRTW

**Don't reinvent the weel**

Use wherever possible existing code, from you or from others.

```
library(someonewrotethatforyou)
```

## RTFM

**Read The Fucking Manual**

Authors of Open Source software have good intentions when they write software for you, but they also make it out of fun. Believe me, if someone writing documentation, fun is not the reason. Only pure good intention with the user is.

## Be consistent

Even if you don't follow any advice here at least be consistent, having your own style is ok, but be consistent

be consistent

be consistent

be consistent

be consistent

be consistent

be consistent

be consistent

be consistent

# Additional Material

## Tools I use

* tidyverse
* rprojects & here package
* git
* knitr
* rmarkdown

## Reference Material

* Reference for the tidyverse [R for Data Science](http://r4ds.had.co.nz)
* Git for R Pogrammer (work in progress) [Happy Git with R](http://happygitwithr.com/classroom-overview.html)
* [much more](https://github.com/aaronpeikert/workshop-reproducible-research/blob/master/05further-reading.md) 

## For Masochists

* very light version: [checkpoint](https://mran.microsoft.com/package/checkpoint) (try that before anything else)
* light Version: [packrat](https://rstudio.github.io/packrat/)
* heavy Version: [docker](https://www.rstudio.com/wp-content/uploads/2018/03/RStudio_Docker_3-9-2018.pdf)
